### 2021-12-08 input/proposal.tex
fs fb+

Algorithm~\ref{algo:add} shows the few instructions that implement a
\emph{reactive} forwarding of best to reach consistent partitioning in
a static network where \processes never join, crash, nor leave the
system. As soon as a \process receives a better message, it delivers
and forwards it. The trade-off consists in decreasing termination time
at cost of increased traffic. For the sake of simplicity and to remove
obvious loops, a \process does not forward a notification to its
sender (Line~\ref{line:accumulator}). Figure~\ref{fig:add} illustrates
the behavior of this algorithm on a system comprising 4 \processes
$a$, $b$, $c$, and $d$. Both $a$ and $d$ become sources.  They \NAMEB
notifications $\alpha_a^0$ and $\alpha_d^0$. They initialize their own
state with the lowest value $0$ (see Line~\ref{line:lowestbound}), and
send a message to each of their neighbors by accumulating the
corresponding edge weight (see Line~\ref{line:accumulator}). In
Figure~\ref{fig:addC}, $b$ receives $\alpha_{d}^{1}$. Since it
improves its own partition distance, it keeps it and forwards it to
its neighbors. In Figure~\ref{fig:addD}, $b$ discards
$\alpha_{a}^{2}$, for it does not improve its partition
distance. Neither $c$ nor $d$ will ever acknowledge that Source~$a$
exists. Ultimately, \processes discard last transiting
messages.
% Despite the obvious lack of traffic optimization,
The system reaches consistent partitioning.

@@

%% At any time, a \process can become a source, hence adding a new
%% partition to the system. This partition eventually includes all
%% \processes that are closer from this new source than any other
%% else. \Processes naturally converge towards their respective best
%% partition by only piggybacking a monotonically increasing distance in
%% forwarded messages.

@@

Delete operations bring a new notion of order between events, and most
importantly between message deliveries. When a \process $s$ performs a
\texttt{Del} operation, this implicitly states that all its preceding
\texttt{Add} operations become obsolete. As consequence, their results
-- scoped broadcast in the form of \texttt{\underline{A}dd}
notifications $\alpha_s$ -- should be canceled by the corresponding
\texttt{\underline{D}el} notifications $\delta_s$.

@@

\begin{definition}[\label{def:lwo}Stale messages]
  Only the latest broadcast of a \node matters.  A message $m$ conveys
  \underline{s}tale control information $\mathcal{S}(m)$ as soon as
  its broadcaster broadcasts another message: $\mathcal{S}(\alpha_x) =
  \exists \delta_x: b_x(\alpha_x) \rightarrow b_x(\delta_x)$.  A
  \process only sends messages that it assumes up-to-date:
  $s_x(\alpha_s) \implies \delta_s \not\in R_x$. \TODO{Only keep last?
    Tothink what happens if we keep all as in def now.}
\end{definition}

A naive implementation based on scoped broadcast as stated in
Definition~\ref{def:scoped} resembles echos in acoustics: staleness
notifications propagate as long as receivers' current partition is the
targeted one, and when staleness notifications reach the bordering
\processes of the deleted partition, it triggers a competition -- or
an echo -- that goes backward to fill the gap left open by removals.

@@

\begin{definition}[\label{def:fs}\underline{F}orwarding of
    \underline{s}taleness (FS)]
  %
  To \NAMEB staleness, a \process $x$ delivers and forwards a received
  staleness notification if it is not yet delivered and targets its
  best delivered message: $\delta_s \in R_x \setminus D_x \wedge
  \alpha_s = \min D_x \TODO{\implies} d_x(\delta_s) \wedge
  f_x(\delta_s)$.
\end{definition}

\begin{definition}[\label{def:fbplus}$\textup{FB}^+$: echos] FB where a
  \process $x$ that receives a staleness notification sends back --~or
  echoes~-- its best up-to-date delivered message: $r_{xy}(\alpha_s)
  \wedge \delta_s \in D_x \implies s_{xy}(\alpha_{s'})$ with
  $\alpha_{s'} = \min D_x \wedge \delta_{s'} \not\in
  D_x$. \TODO{Rework min}
\end{definition}

Figure~\ref{fig:del} illustrates the behavior of this intuitive
implementation.  Two partitions initially exist: $P_a$ and $P_d$ that
respectively include $\{a\}$, and $\{b, c, d\}$. In
Figure~\ref{fig:delA}, $a$ deletes its partition. It notifies all
neighboring \processes~--~here only $b$~-- that may belong to its
partition. Upon receipt, $b$ discards the delete notification
$\delta_a$ since $\delta_a$ does not target the former's partition
$P_d$. \Process~$b$ sends its own best partition $\alpha_d^3$ that may
be the best for $a$. Eventually, every \process belongs to Partition
$P_d$. In this scenario, they reach consistent
partitioning. Unfortunately, this protocol does not ensure consistent
partitioning in every scenario.

>>>>

Algorithm~\ref{algo:add} shows the few instructions that implement a
\emph{reactive} forwarding of best to reach consistent partitioning in
a static network where \processes never join, crash, nor leave the
system. As soon as a \process receives a better message, it delivers
and forwards it. The trade-off consists in decreasing termination time
at cost of increased traffic. For the sake of simplicity and to remove
obvious loops, a \process does not forward a notification to its
sender (Line~\ref{line:accumulator}). Figure~\ref{fig:add} illustrates
the behavior of this algorithm on a system comprising 4 \processes
$a$, $b$, $c$, and $d$. Both $a$ and $d$ become sources.  They \NAMEB
notifications $\alpha_a^0$ and $\alpha_d^0$. They initialize their own
state with the lowest value $0$ (see Line~\ref{line:lowestbound}), and
send a message to each of their neighbors by accumulating the
corresponding edge weight (see Line~\ref{line:accumulator}). In
Figure~\ref{fig:addC}, $b$ receives $\alpha_{d}^{1}$. Since it
improves its own partition distance, it keeps it and forwards it to
its neighbors. In Figure~\ref{fig:addD}, $b$ discards
$\alpha_{a}^{2}$, for it does not improve its partition
distance. Neither $c$ nor $d$ will ever acknowledge that Source~$a$
exists. Ultimately, \processes discard last transiting
messages.
% Despite the obvious lack of traffic optimization,
The system reaches consistent partitioning.
\TODO{Traffic complexity here?}

@@

@@

Delete operations bring a new notion of order between events, and most
importantly between message deliveries. A \process $s$ that performs a
\texttt{Del} operation implicitly states that all its preceding
\texttt{Add} operations become obsolete. As consequence, their results
-- scoped broadcast in the form of \texttt{\underline{A}dd}
notifications $\alpha_s$ -- should be canceled by the corresponding
\texttt{\underline{D}el} notifications $\delta_s$.

@@

\begin{definition}[\label{def:lwo}Stale messages]
  Only the latest broadcast of a \node matters.  A message $m$ conveys
  \underline{s}tale control information $\mathcal{S}(m)$ as soon as
  its broadcaster broadcasts another message: $\mathcal{S}(\alpha_x) =
  \exists \delta_x: b_x(\alpha_x) \rightarrow b_x(\delta_x)$.  A
  \process only sends messages that it assumes
  \underline{u}p-\underline{t}o-\underline{d}ate.
  % : $s_x(\alpha_s)
  % \implies \delta_s \not\in R_x$.
  For convenience, we define $\textup{minUTD(}D\textup{)} = \min
  \{\alpha_s^d | \alpha_s^d \in D \wedge \delta_s \notin D\}$.
  %  \TODO{Only keep last?  Tothink what happens if we keep all as in
%    def now.}
\end{definition}

A naive attempt at implementing DCP resembles echoes in
acoustics. Following the principles of scoped broadcast as stated in
Definition~\ref{def:scoped}, staleness notifications propagate as long
as receivers' current partition is the targeted one, and when
staleness notifications reach the bordering \processes of the deleted
partition, it triggers a competition -- or an echo -- that goes
backward to fill the gap left open by removals.

@@

\begin{definition}[\label{def:fs}\underline{F}orwarding of
    \underline{s}taleness (FS)]
  %
  A \process $x$ delivers and forwards a received staleness
  notification if it targets its best up-to-date delivered message:
  $(\delta_s \in R_x \wedge \alpha_s = \textup{minUTD}~D_x) \iff
  \eventually (d_x(\delta_s) \rightarrow f_x(\delta_s))$.
\end{definition}

\begin{definition}[\label{def:fbplus}$\textup{FB}^+$: echos] FB where a
  \process $x$ that receives a staleness notification sends back --~or
  echoes~-- its best up-to-date delivered message: $r_{xy}(\alpha_s)
  \wedge \delta_s \in D_x \implies s_{xy}(\alpha_{s'}^d\oplus
  ^{w_{xy}})$ with $\alpha_{s'^d} = \min D_x \wedge \delta_{s'} \not\in
  D_x$. \TODO{Rework min}
\end{definition}

Figure~\ref{fig:del} illustrates the behavior of this implementation
($\textup{FB}^+ \wedge \textup{FS}$).  Two partitions initially exist:
$P_a$ and $P_d$ that respectively include $\{a\}$, and $\{b, c,
d\}$. In Figure~\ref{fig:delA}, $a$ deletes its partition. FS in
Definition~\ref{def:fs} states that $a$ must notify all its
neighbors~--~here only $b$~-- that may belong to its partition.  In
Figure~\ref{fig:delB}, $b$ receives but does not deliver $\delta_a$
since $\delta_a$ does not target its current partition $P_d$.
$\textup{FB}^+$ in Definition~\ref{def:fbplus} states that $b$ must
echo back its own best up-to-date message $\alpha_d^3$, for it may be
the best for $a$. Eventually, every \process belongs to Partition
$P_d$. In this scenario, they reach consistent
partitioning. Unfortunately, this protocol does not ensure consistent
partitioning in every scenario.
